---
title: "Airbnb"
author: "Rodrigo Vecio"
output: html_notebook
---

## Packages
```{r}
library(tidyverse)
library(dsbox)
```

## Load and glimpse
The data is in the dsbox package and it’s called edibnb.
You can view the data using
```{r}
glimpse(edibnb)
```

## Question 1: Number of observations
How many Airbnb listings are included in this dataset? Use the following code block to find out.
```{r}
nrow(edibnb)
```

## Question 2: Examining further
Each column represents a variable. We can get a list of the variables in the data frame using the names() function.
```{r}
names(edibnb)
```

## Question 3: Building a histogram
Create a histogram of the distribution of Airbnb listing prices (nightly rates) in Edinburgh. Sample code is provided below, but you will need to fill in the blanks.
```{r}
ggplot(data = edibnb, mapping = aes(x = price)) +
  geom_histogram() +
  labs(
    x = "Airbnb listing price, in £",
    y = "Frequency",
    title = "Distribution of Airbnb nightly rates in Edinburgh"
  )
```

## Question 4: Faceted histograms
Create a faceted histogram where each facet represents a neighbourhood and displays the distribution of Airbnb prices in that neighbourhood. Sample code is provided below, but you will need to fill in the blanks.
```{r}
ggplot(data = edibnb, mapping = aes(x = price)) +
  geom_histogram() +
  facet_wrap(~neighbourhood) +
  labs(
    x = "Airbnb listing price, in £",
    y = "Frequency",
    title = "Distribution of Airbnb nightly rates in Edinburgh",
    subtitle = "By neighbourhood"
  )
```

## Question 5: Binwidth experimentation
Instead of having R pick a binwidth itself, try out different binwidths using the code below:
```{r}
ggplot(data = edibnb, mapping = aes(x = price)) +
  geom_histogram(binwidth = 75) +
  facet_wrap(~neighbourhood) +
  labs(
    x = "Airbnb listing price, in £",
    y = "Frequency",
    title = "Distribution of Airbnb nightly rates in Edinburgh"
  )
```

## Question 6: Visualising review scores
Create a similar visualisation, this time showing the distribution of review scores (review_scores_rating) across neighbourhoods.
```{r}
ggplot(data = edibnb, mapping = aes(x = review_scores_rating)) +
  geom_histogram() +
  facet_wrap(~neighbourhood) +
  labs(
    x = "Average rating score of property",
    y = "Frequency",
    title = "Distribution of Airbnb rating scores in Edinburgh",
    subtitle = "By neighbourhood"
  )
```