---
title: "Lego Sales"
author: "Rodrigo Vecino"
output: html_notebook
---

## Packages
```{r}
library(tidyverse)
library(dsbox)
```

## Quesrtion 1: Most common names
What are the three most common first names of purchasers? Use the code chunk below to find out!
```{r}
lego_sales %>%
  count(first_name) %>%
  arrange(desc(n))
```

## Quesrtion 2: Most common themes
What are the four most common themes of lego sets purchased? Use the code chunk below to find out!
```{r}
lego_sales %>%
  count(theme) %>%
  arrange(desc(n))
```

## Quesrtion 3: Most common subtheme
Among the most common theme of lego sets purchased, what is the most common subtheme?
```{r}
lego_sales %>%
  filter(theme == "Star Wars") %>%
  count(subtheme) %>%
  arrange(desc(n))
```

## Quesrtion 4: Grouping by age
Create a new variable called age_group and group the ages of customers into the following categories: “18 and under”, “19 - 25”, “26 - 35”, “36 - 50”, “51 and over”.
```{r}
lego_sales <- lego_sales %>%
  mutate(age_group = case_when(
    age <= 18 ~ "18 and under",
    age >= 19 & age <= 25 ~ "19 - 25", 
    age >= 26 & age <= 35 ~ "26 - 35", 
    age >= 36 & age <= 50 ~ "36 - 50", 
    age >= 51 ~ "51 and over" 
  ))
```

## Quesrtion 5: Most common age group
Count the number of customers in each age group and display the counts in descending order, from most common to least common age group.
Which age group is the most common? Write code in the chunk below to figure it out!
```{r}
lego_sales %>%
  count(age_group) %>%
  arrange(desc(n))
```

## Quesrtion 6: Quesrtion : Quantity of purchases by age
Which age group has purchased the highest number of lego sets?
```{r}
lego_sales %>%
  group_by(age_group) %>%
  summarize(
    total_n = sum(quantity)
  ) %>%
  arrange(desc(total_n))
```

## Quesrtion 7: Most generous age group
Which age group has spent the most money on legos?
```{r}
lego_sales %>%
  mutate(
    amount_spent = us_price * quantity
  ) %>%
  group_by(age_group) %>%
  summarize(
    total_spent = sum(amount_spent)
  ) %>%
  arrange(desc(total_spent))
```