---
title: "Books"
output: html_notebook
author : "Rodrigo Vecino"
---

## Packages
```{r packages}
library(dplyr)
library(janitor)
library(tidyverse)
```

## Load data
```{r data}
 books = read.csv(file="https://raw.githubusercontent.com/jskufca/DS241_bikes_2021/Vecino/Portfolio/Data/CU_ACAD_DEPT_WHEN_OFF_ENRL_22740.csv", header=TRUE, skip=1)
```

## Variable Calculus
```{r calculus}
title <-c("Calculus I", "Calculus II")
q <- c("LEC")

calculus <- books %>%
  filter(Title %in% title) %>%
  filter(Component %in% q) %>%
  group_by(Term.1) %>%
  arrange(Term) %>%
  ungroup()
```

## Variable Fall_calc1
Create Variable with all the students from the fall semester in Calculus 1.
```{r fall_calc1}
term_fall <-seq(4149, 4219, by = 10)

fall_calc1 <- calculus %>%
  filter(calculus$Term %in% term_fall) %>%
  group_by(Term.1)

fall_calc1 = filter(fall_calc1, Title == "Calculus I")

fall_calc1 = subset(fall_calc1, select = -c(Acad.Org, Current.Status, Course.ID, Subject, Instructor, Component, Catalog, Section, Term, Title) )

fall_calc1 <- fall_calc1 %>%
  summarise(enroll = sum(Tot.Enrl)) %>%
  group_by(Term.1)

names(fall_calc1)[names(fall_calc1) == 'enroll'] <- 'fall_calc1'
```

## Variable Fall_calc2
Create Variable with all the students from the fall semester in Calculus 2.
```{r fall_calc2}
term_fall <-seq(4149, 4219, by = 10)

fall_calc2 <- calculus %>%
  filter(calculus$Term %in% term_fall) %>%
  group_by(Term.1)

fall_calc2 = filter(fall_calc2, Title == "Calculus II")

fall_calc2 = subset(fall_calc2, select = -c(Acad.Org, Current.Status, Course.ID, Subject, Instructor, Component, Catalog, Section, Term, Title) )

fall_calc2 <- fall_calc2 %>%
  summarise(enroll = sum(Tot.Enrl)) %>%
  group_by(Term.1)

names(fall_calc2)[names(fall_calc2) == 'enroll'] <- 'fall_calc2'
```

## Variable Spring_calc1
Create Variable with all the students from the spring semester in Calculus 1.
```{r spring_calc1}
term_spring <-seq(4152, 4222, by = 10)

spring_calc1 <- calculus %>%
  filter(calculus$Term %in% term_spring) %>%
  group_by(Term.1)

spring_calc1 = filter(spring_calc1, Title == "Calculus I")

spring_calc1 = subset(spring_calc1, select = -c(Acad.Org, Current.Status, Course.ID, Subject, Instructor, Component, Catalog, Section, Term, Title) )

spring_calc1 <- spring_calc1 %>%
  summarise(enroll = sum(Tot.Enrl)) %>%
  group_by(Term.1)

names(spring_calc1)[names(spring_calc1) == 'enroll'] <- 'spring_calc1'
```

## Variable Spring_calc2
Create Variable with all the students from the spring semester in Calculus 2.
```{r spring_calc2}
term_spring <-seq(4152, 4222, by = 10)

spring_calc2 <- calculus %>%
  filter(calculus$Term %in% term_spring) %>%
  group_by(Term.1)

spring_calc2 = filter(spring_calc2, Title == "Calculus II")

spring_calc2 = subset(spring_calc2, select = -c(Acad.Org, Current.Status, Course.ID, Subject, Instructor, Component, Catalog, Section, Term, Title) )

spring_calc2 <- spring_calc2 %>%
  summarise(enroll = sum(Tot.Enrl)) %>%
  group_by(Term.1)

names(spring_calc2)[names(spring_calc2) == 'enroll'] <- 'spring_calc2'
```


## Prediction
Create a model to predict the number of students in calculus 2 the spring semester.
```{r prediction - other way}
z = fall_calc1[1:7, 2]
z[2] = fall_calc2[1:7, 2]
z[3] = spring_calc1[1:7, 2]
z[4] = spring_calc2[1:7, 2]

new = fall_calc1[8, 2]
new[2] = fall_calc2[8, 2]

install.packages("caret")
library(caret)

model <- train(spring_calc2  ~ fall_calc1 + fall_calc2, data = z, method="lm")
predictions <- predict(model, new, interval = "prediction")
final <- cbind(z, predictions)

ggplot(final, aes(spring_calc2, fall_calc1)) +
geom_point() +
stat_smooth(method = lm)
```



