---
title: "COVID IN US STATES - CODE CDC"
author: "Rodrigo Vecino"
output:
  html_document:
    df_print: paged
---

data from: https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv

## Packages
```{r load-packages, warning=FALSE, message=FALSE}
# devtools::install_github("RamiKrispin/coronavirus")
library(coronavirus)
library(tidyverse)
library(lubridate)
library(glue)
library(scales)
library(ggrepel)
library(DT)
library(dplyr)
library(data.table)
```

## DATA US STATES
Load the data.
```{r load data states}

data_states <- read.csv(file = "https://raw.githubusercontent.com/jskufca/DS241_bikes_2021/Vecino/Portfolio/Data/data_CDC.csv", header = TRUE)
```


## Data prep
I create a variable with all the states that I am interested in.
```{r select-states}
states <- c(
  "FL",
  "NY",
  "NJ",
  "CT",
  "MSS",
  "TX",
  "CAL",
  "WASH",
  "IL",
  "ID"
)
```

I filter only data with the states I am interested in and I record the days elapsed, end date and lebels.
```{r prep-state-data}
states_deaths <- data_states %>%
  filter( state %in% states) %>%
  arrange(date) %>%
  ungroup()

states_deaths$date <- as.Date(states_deaths$date, "%m/%d/%Y")

days_elapsed <- as.numeric(states_deaths$date - min(states_deaths$date))
end_date     <- if_else(states_deaths$date == max(states_deaths$date), TRUE, FALSE)
end_label    <- if_else(end_date, states_deaths$state, NULL)

states_deaths$days_elapsed = days_elapsed
states_deaths$end_date = end_date
states_deaths$end_label = end_label
```

I also need to take a note of the "as of date" for the data so that we can properly label our visualization.
```{r record-as-of-date}
as_of_date <- states_deaths %>% 
  summarise(max(date)) %>% 
  pull()

as_of_date_formatted <- glue("{wday(as_of_date)}, {month(as_of_date)} {day(as_of_date)}, {year(as_of_date)}")
```

## Visualization - Deaths per day from COVID-19
```{r visualise, warning=FALSE}
ggplot(data = states_deaths,
       mapping = aes(x = date, 
                     y = new_death, 
                     color = state, 
                     label = end_label)) +
  # represent cumulative cases with lines
  geom_line(size = 0.7, alpha = 0.8) +
  # add points to line endings
  geom_point(data = states_deaths %>% filter(end_date)) +
  # add country labels, nudged above the lines
  geom_label_repel(nudge_y = 1, direction = "y", hjust = 1) + 
  # turn off legend
  guides(color = FALSE) +
  # use pretty colors
  scale_color_viridis_d() +
  # better formatting for y-axis
  scale_y_continuous(labels = label_comma()) +
  # use minimal theme
  theme_minimal() +
  # customize labels
  labs(
    x = "Date",
    y = "Deaths",
    title = "Deaths per day from COVID-19, selected states",
    subtitle = glue("Data as of", as_of_date_formatted, .sep = " ")
  )
```

## Visualization - Cumulative deaths from COVID-19
```{r visualise 2, warning=FALSE}
ggplot(data = states_deaths,
       mapping = aes(x = days_elapsed, 
                     y = tot_death, 
                     color = state, 
                     label = end_label)) +
  # represent cumulative cases with lines
  geom_line(size = 0.7, alpha = 0.8) +
  # add points to line endings
  geom_point(data = states_deaths %>% filter(end_date)) +
  # add country labels, nudged above the lines
  geom_label_repel(nudge_y = 1, direction = "y", hjust = 1) + 
  # turn off legend
  guides(color = FALSE) +
  # use pretty colors
  scale_color_viridis_d() +
  # better formatting for y-axis
  scale_y_continuous(labels = label_comma()) +
  # use minimal theme
  theme_minimal() +
  # customize labels
  labs(
    x = "Days since 10th confirmed death",
    y = "Cumulative number of deaths",
    title = "Cumulative deaths from COVID-19, selected states",
    subtitle = glue("Data as of", as_of_date_formatted, .sep = " ")
  )
```

## Visualization - Cases from COVID-19
```{r visualise 3, warning=FALSE}
ggplot(data = states_deaths,
       mapping = aes(x = days_elapsed, 
                     y = new_case, 
                     color = state, 
                     label = end_label)) +
  # represent cumulative cases with lines
  geom_point(size = 0.7, alpha = 0.8) +
  # add country labels, nudged above the lines
  geom_label_repel(nudge_y = 1, direction = "y", hjust = 1) + 
  # turn off legend
  guides(color = FALSE) +
  # use pretty colors
  scale_color_viridis_d() +
  # better formatting for y-axis
  scale_y_continuous(labels = label_comma()) +
  # use minimal theme
  theme_minimal() +
  # customize labels
  labs(
    x = "Days since 10th confirmed death",
    y = "Cases",
    title = "Cases from COVID-19, selected countries",
    subtitle = glue("Data as of", as_of_date_formatted, .sep = " "),
    caption = "Source: github.com/RamiKrispin/coronavirus"
  )
```