---
title: "Lab04"
author: "Rodrigo Vecino"
date: "10/08/2021"
output: html_notebook
---


```{r packages}
library(tidyverse) 
devtools::install_github("rstudio-education/dsbox")
library(dsbox) 
```

```{r data}
states <- read.csv(url("https://raw.githubusercontent.com/rstudio-education/datascience-box/master/course-materials/starters/lab/lab-05-wrangle-sp-data/data/states.csv"))
```

## Exercise 1
What are the dimensions of the Denny’s dataset?
(Hint: Use inline R code and functions like nrow and ncol to compose your answer.)
What does each row in the dataset represent?
What are the variables?

```{r exercise1}
nrow(dennys)
# 1643
view(dennys)
# the location of each restaurant
names(dennys)
# address  city state zip longitude latitude
```


## Exercise 2
What are the dimensions of the La Quinta’s dataset?
What does each row in the dataset represent?
What are the variables?

```{r exercise2}
nrow(laquinta)
# 909
view(laquinta)
# the location of each restaurant
names(laquinta)
# address  city state zip longitude latitude
```

## Exercise 3

Take a look at the websites that the data come from (linked above).
Are there any La Quinta’s locations outside of the US?
If so, which countries?
What about Denny’s?
```{r exercise3}
# dennys: 0
# laquinta: 14
```

## Exercise 4

Now take a look at the data.
What would be some ways of determining whether or not either establishment has any locations outside the US using just the data (and not the websites).
Don’t worry about whether you know how to implement this, just brainstorm some ideas.
Write down at least one as your answer, but you’re welcomed to write down a few options too.

```{r exercise4}
states_name = subset(states, select = -c(area, name))

new_dennys <- dennys %>%
  filter(!(state %in% states_name$abbreviation))
nrow(new_dennys)
# dennys: 0

new_laquinta <- laquinta %>%
  filter(!(state %in% states_name$abbreviation))
nrow(new_laquinta)
# laquinta: 14
```

## Exercise 5

Find the Denny’s locations that are outside the US, if any.
To do so, filter the Denny’s locations for observations where state is not in states$abbreviation.
The code for this is given below.
Note that the %in% operator matches the states listed in the state variable to those listed in states$abbreviation.
The ! operator means not.
Are there any Denny’s locations outside the US?
```{r exercise5}
dennys %>%
  filter(!(state %in% states$abbreviation))
# dennys: 0
```

## Exercise 6

Add a country variable to the Denny’s dataset and set all observations equal to "United States".
Remember, you can use the mutate function for adding a variable.
Make sure to save the result of this as dennys again so that the stored data frame contains the new variable going forward.
```{r exercise6}
dennys <- dennys%>%
  mutate(country = "United States")
```

## Exercise 7
Find the La Quinta locations that are outside the US, and figure out which country they are in.
This might require some googling.
Take notes, you will need to use this information in the next exercise.
```{r exercise7}
# Chile, Colombia, United Arab Emirates, Turkey, Honduras, New Zealand, Mexico, Canada
```

## Exercise 8
Add a country variable to the La Quinta dataset.
Use the case_when function to populate this variable.
You’ll need to refer to your notes from Exercise 7 about which country the non-US locations are in.
Here is some starter code to get you going:
```{r exercise8}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == "FM"           ~ "Honduras",
    state %in% c("IST", "BXN") ~ "Turkey",
    state == "PQ"           ~ "New Zeland",
    state %in% c("AG", "NL", "QR", "CH", "PU", "SL") ~ "Mexico"
  ))

laquinta <- laquinta %>%
   filter(country == "United States")
```

## Exercise 9
Which states have the most and fewest Denny’s locations?
What about La Quinta?
Is this surprising?
Why or why not?
```{r exercise9}
dennys_n <- dennys%>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(n)
# California: 403
# Delaware: 1

laquinta_n <- laquinta%>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(n)
# Texas: 237
# Maine: 1

# Is not surprising. laquinta and dennys do not have a linear relation
```

## Exercise 10
Which states have the most Denny’s locations per thousand square miles?
What about La Quinta?

```{r exercise10}
dennys_n$npsm = dennys_n$n / dennys_n$area
dennys_n <- dennys_n %>%
  arrange(npsm)
# District of Columbia

laquinta_n$npsm = laquinta_n$n / laquinta_n$area
laquinta_n <- laquinta_n %>%
  arrange(npsm)
# Rhode Island
```

## Exercise 11
Filter the data for observations in North Carolina only, and recreate the plot.
You should also adjust the transparency of the points, by setting the alpha level, so that it’s easier to see the overplotted ones.
Visually, does Mitch Hedberg’s joke appear to hold here?

```{r exercise11}
dennys_1 <- dennys %>%
  mutate(establishment = "Denny's") %>%
  filter(state =="NC")
laquinta_1 <- laquinta %>%
  mutate(establishment = "La Quinta") %>%
  filter(state =="NC")

dn_lq <- bind_rows(dennys_1, laquinta_1)

ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```

## Exercise 12
Now filter the data for observations in Texas only, and recreate the plot, with an appropriate alpha level.
Visually, does Mitch Hedberg’s joke appear to hold here?

```{r exercise12}
dennys_2 <- dennys %>%
  mutate(establishment = "Denny's") %>%
  filter(state =="TX")
laquinta_2 <- laquinta %>%
  mutate(establishment = "La Quinta") %>%
  filter(state =="TX")

dn_lq <- bind_rows(dennys_2, laquinta_2)

ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```