---
title: "Making maps in R:"
subtitle: "Using the sf and tmap Packages"
author: "Rodrigo Vecino"
---


# Libraries

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

library(tidyverse)  # Modern data science workflow
library(sf)         # Simple features for R
library(tmap)       # Thematic Maps
library(tmaptools)  # Thematic Maps Tools
library(RColorBrewer) # ColorBrewer Palettes
library(leaflet)    # Interactive web maps
library(rgdal)      # Bindings for the Geospatial Data Abstraction Library
library(rgeos)      # Interface to Geometry Engine - Open Source 


# Change the presentation of decimal numbers to 4 and avoid scientific notation
options(prompt="R> ", digits=4, scipen=999)
```


# Import the data

## Non-spatial data

```{r}
dat <-read.csv("U:/Data Science/MAP/DATASET.csv")
```

### Explore the data

```{r}
glimpse(dat) 
```

#### Definitions of variables

- growth: growth rate of per capita real income 1987-2001
- base: lograrithm of the per apita income in the base year 1987
- T: average terrorism index
- E: average years of schooling
- G: real per capita government expenditures in 1987
- Coastal: dummy variable  which takes the value of one if the province is a coastal province.

## Spatial data

```{r}
mapData <- read_sf("U:/Data Science/MAP/MAP.shp")
```

### Explore the data

```{r}
glimpse(mapData)
```

- Check the Coordinate Reference System

```{r}
st_crs(mapData)
```


# Transform the data

No need to transform any data because both datasets share a common variable `id`


# Merge the data


```{r}
dat_map <- inner_join(
  dat,
  mapData,
  by = "id"
)
```

## Keep data as sf object

- Keep the data as sf class, so we will not lose the coodinate system

```{r}
dat_map <- st_as_sf(dat_map)
st_crs(mapData)
```

```{r}
glimpse(dat_map)
```



# Plot Thematic Maps


## Quick Map 

- A quick map of the Terrorism variable

```{r}
qtm(dat_map, fill = "T")
```


## Simple Map

```{r}
tm_shape(dat_map) + tm_fill("T") 
```


## Remove frame

```{r}
tm_shape(dat_map) + 
  tm_fill("T") +
  tm_layout(frame = FALSE)
```


## Set legend outside

```{r}
tm_shape(dat_map) + tm_fill("T") +
  tm_layout(legend.outside = TRUE, frame = FALSE)
```

## Set legend size

```{r}
tm_shape(dat_map) + tm_fill("T") +
  tm_layout(legend.outside = TRUE, frame = FALSE) +
  tm_legend(text.size = 0.3) 
```



## Set color palette

See color reference [here](https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/)

```{r}
tm_shape(dat_map) + 
  tm_fill("T", palette = "Greens") +
  tm_layout(legend.outside = TRUE, frame = FALSE)
```

```{r}
tm_shape(dat_map) + 
  tm_fill("T", palette = "-Greens") +
  tm_layout(legend.outside = TRUE, frame = FALSE)
```


```{r}
tm_shape(dat_map) + 
  tm_fill("T", palette = "viridis") +
  tm_layout(legend.outside = TRUE, frame = FALSE)
```


## Add polygon names

```{r}
tm_shape(dat_map) + 
  tm_fill("T", palette = "viridis") +
  tm_layout(legend.outside = TRUE, frame = FALSE) +
  tm_text("province", size = "Shape_Area",  auto.placement = F, legend.size.show = FALSE) 
```



[Documentation](https://rdrr.io/cran/tmap/man/tm_text.html)

## Add scale bar

```{r}
tm_shape(dat_map) + 
  tm_fill("T", palette = "viridis") +
  tm_layout(legend.outside = TRUE, frame = FALSE) +
  tm_text("province", size = "Shape_Area",  auto.placement = F, legend.size.show = FALSE)  + 
  tm_scale_bar(position = c("RIGHT", "BOTTOM"))
```


## Set color intervals

Enter “style =” followed by one of the options below.

- equal: divides the range of the variable into n parts.
- pretty: chooses a number of breaks to fit a sequence of equality spaced ‘round’ values. 
- quantile: equal number of cases in each group
- jenks: looks for natural breaks in the data
- cat: when the variable is categorical 

```{r}
tm_shape(dat_map) + 
  tm_fill("T",
          style = "quantile",
          palette = "Reds"
    ) +
  tm_layout(
    legend.outside = TRUE,
    frame = FALSE)
```

Change the number of intervals in the color scheme and how the intervals are spaced. Changing the number of intervals n = 7. So, we have 7 shades instead of the default 5.

```{r}
tm_shape(dat_map) + 
  tm_fill("T",
          style = "quantile",
          n = 7,
          palette = "Reds"
    ) +
  tm_layout(
    legend.outside = TRUE,
    frame = FALSE)
```



## Add histogram

```{r}
tm_shape(dat_map) + 
  tm_fill("T",
          style = "quantile",
          n = 5,
          palette = "YlOrBr",
          legend.hist = TRUE
    ) +
  tm_layout(
    legend.outside = TRUE,
    frame = FALSE)
```


## Add borders

You can edit the borders of the shapefile with the tm_borders() function which has many arguments. alpha denotes the level of transparency on a scale from 0 to 1 where 0 is completely transparent.

```{r}
tm_shape(dat_map) + 
  tm_fill("T",
          style = "quantile",
          palette = "Blues"
    ) + 
  tm_borders(alpha=.4) +
  tm_layout(
    legend.outside = TRUE,
    frame = FALSE)
```

## Add compass

```{r}
tm_shape(dat_map) + 
  tm_fill("T",
          style = "quantile",
          palette = "viridis"
    ) + 
  tm_borders(alpha=.4) + 
  tm_compass(type = "8star", 
              position = c("RIGHT", "BOTTOM"),
              show.labels = 2,
              text.size = 0.35) +
  tm_layout(legend.outside = TRUE,
            frame = FALSE)
```

[Documentation](https://www.rdocumentation.org/packages/tmap/versions/2.3-2/topics/tm_compass)


## Edit the layout

```{r}
tm_shape(dat_map) + 
  tm_fill("T",
          palette = "viridis",
          style = "quantile",
          title = "Terrorist Attacks"
          ) + 
tm_borders(alpha=.4) + 
tm_layout(
  legend.text.size = 0.7,
  legend.title.size = 1,
  legend.position = c("right", "bottom"),
  legend.outside = TRUE,
  frame = FALSE
  ) 
```

[Documentation](https://www.rdocumentation.org/packages/tmap/versions/2.3-2/topics/tm_layout)

## Interactive map


```{r}
tmap_mode("view")

tm_shape(dat_map) + 
  tm_fill("T",
          palette = "viridis",
          id="province",
          popup.vars=c("T", "growth", "base", "E", "G")
          ) + 
	tm_legend(outside=TRUE) +
  tm_layout(frame = FALSE) 

tmap_mode("plot")
```


[Documentation](https://rdrr.io/cran/tmap/man/tm_view.html)

# Save a new shapefile

```{r eval=FALSE}
st_write(dat_map, "dat_map.shp")
```

